---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
Christian Willig  
August 2015

## Loading and preprocessing the data
In order to process the data I downloaded the file and load the data into a dataframe `data`.
```{r}
#creating temporary location for the file to be downloaded
temp <- tempfile()
urlfile <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

#Downloading the file
download.file(urlfile,temp, mode="wb")

#Un compressing the file into the temporary location
unzip(temp)
unlink(temp)

#creating dataframe with the activity data
data <- read.csv("activity.csv")

#making sure date behaves as a date
data <- transform(data, date = as.Date(date))
```


## What is the mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.  
1. Make a histogram of the total number of steps taken each day  

```{r}
#ignoring missing values
data_df2 <- na.omit(data)

#Calculate total steps during a day
total_steps_day <- aggregate(steps ~ date, data_df2, sum)

#Draw histogram with the total steps during a day
hist(total_steps_day$steps, main = paste("Total Steps per Day"), col="blue", xlab="Number of Steps")

```

2. Calculate and report the mean and median total number of steps taken
per day  

```{r}
#Calculate median and mean
mean_steps_per_day <- mean(total_steps_day$steps)
median_steps_per_day <- median(total_steps_day$steps)

```

The `mean` of total number of steps per day is `r format(mean_steps_per_day, sciencific = T)`

The `median` of total number of steps per day is `r format(median_steps_per_day, sciencific = T)`

## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all days (y-axis)  

```{r}
steps_by_interval <- aggregate(steps ~ interval, data_df2, mean)

plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Avg Number of Steps per Day by Interval")

```

2. Which 5-minute interval, on average across all the days in the dataset,
contains the maximum number of steps?  

```{r}
max_number_of_steps <- steps_by_interval[which.max(steps_by_interval$steps),1]

```
The 5-minute interval that contains the maximum number of steps is `r format(max_number_of_steps, sciencific = T)`

## Imputing missing values
Note that there are a number of days/intervals where there are missing values
(coded as NA). The presence of missing days may introduce bias into some
calculations or summaries of the data.  

1. Calculate and report the total number of missing values in the dataset
(i.e. the total number of rows with NAs)  

```{r}
#Counting missing values from the dataset.
number_of_na <- sum(is.na(data$steps))

```

There are `r format(number_of_na, sciencific = T)` missing values in the dataset.

2. Devise a strategy for filling in all of the missing values in the dataset. The
strategy does not need to be sophisticated. For example, you could use
the mean/median for that day, or the mean for that 5-minute interval, etc.  

I chose to use the mean for that 5-minute interval. This means that NA values will be replace by the mean of that 5-minute interval.

In order to accomplish this we process all NA through a loop.

3. Create a new dataset that is equal to the original dataset but with the
missing data filled in. 

```{r}
#Performing imputation of NAs.
for (i in 1:nrow(data)){
  if (is.na(data$steps[i])){
    interval_val <- data$interval[i]
    row_id <- which(steps_by_interval$interval == interval_val)
    steps_val <- steps_by_interval$steps[row_id]
    data$steps[i] <- steps_val
  }
}

```


4. Make a histogram of the total number of steps taken each day and Calculate
and report the mean and median total number of steps taken per day. Do
these values differ from the estimates from the first part of the assignment?
What is the impact of imputing missing data on the estimates of the total
daily number of steps?  

```{r}
steps_by_interval_imputed <- aggregate(steps ~ date, data, sum)

hist(steps_by_interval_imputed$steps, col=1, main="(Imputed) Histogram of total number of steps per day", xlab="Total number of steps in a day")

#Calculate the mean and median of total number of steps per day
imputed_mean <- mean(steps_by_interval_imputed$steps)

imputed_median <- median(steps_by_interval_imputed$steps)
```

The new `mean` and `median` for the imputed dataset are `r format(imputed_mean, sciencific = T)` and `r format(imputed_median, sciencific = T)` respectivitely.

Compared to the original values calculated for the initial dataset, there is no difference in the mean whereas there is a slight difference in the median value.


## Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.  

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")

#Adding column to dataset with no missing values.
data$day_of_week = as.factor(ifelse(is.element(weekdays(as.Date(data$date)),weekdays), "Weekday", "Weekend"))

```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).  

```{r}
steps_by_interval_2 <- aggregate(steps ~ interval + day_of_week, data, mean)

library(lattice)

xyplot(steps_by_interval_2$steps ~ steps_by_interval_2$interval|steps_by_interval_2$day_of_week, main="Avg Steps per Day by Interval",xlab="Intervals", ylab="Steps",layout=c(1,2), type="l")

```

