---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
Christian Willig  
August 2015

## Loading and preprocessing the data
In order to process the data I downloaded the file and load the data into a dataframe `data`.
```{r}
#creating temporary location for the file to be downloaded
temp <- tempfile()
urlfile <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

#Downloading the file
download.file(urlfile,temp, mode="wb")

#Un compressing the file into the temporary location
unzip(temp)
unlink(temp)

#creating dataframe with the activity data
data <- read.csv("activity.csv")

#making sure date behaves as a date
data <- transform(data, date = as.Date(date))
```


## What is the mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.  
1. Make a histogram of the total number of steps taken each day  
2. Calculate and report the mean and median total number of steps taken
per day  

```{r}
#ignoring missing values
data_df2 <- na.omit(data)

#Calculate total steps during a day
total_steps_day <- aggregate(steps ~ date, data_df2, sum)

#Draw histogram with the total steps during a day
hist(total_steps_day$steps, main = paste("Total Steps per Day"), col="blue", xlab="Number of Steps")

#Calculate median and mean
mean_steps_per_day <- mean(total_steps_day$steps)
median_steps_per_day <- median(total_steps_day$steps)
```

The `mean` of total number of steps per day is `r format(mean_steps_per_day, sciencific = T)`

The `median` of total number of steps per day is `r format(median_steps_per_day, sciencific = T)`

## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all days (y-axis)  

```{r}
steps_by_interval <- aggregate(steps ~ interval, data_df2, mean)

plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Avg Number of Steps per Day by Interval")

```

2. Which 5-minute interval, on average across all the days in the dataset,
contains the maximum number of steps?  

```{r}
max_number_of_steps <- steps_by_interval[which.max(steps_by_interval$steps),1]

```
The 5-minute interval that contains the maximum number of steps is `r format(max_number_of_steps, sciencific = T)`

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
